The Way Forward in Parallelizing Dynamic Languages

### 2014-05-03 talk/icooolps2014/position-paper.tex
Add a suggestion
Precision

In the end, fine-grained locking can transparently replace the GIL
and therefore parallelise existing applications without any
changes. It does however not provide a better synchronisation
mechanism to the application like e.g. atomic blocks.

@@

While one can argue that STM requires the insertion of read and write
barriers in the whole program, this can be done automatically and
locally by a program transformation\cite{felber07}. There are attempts
to do the same for fine-grained locking\cite{bill06} but they require
a whole program analysis since locks are inherently non-composable.
The effectiveness of these approaches still has to be proven for our
use case.

>>>>

In the end, fine-grained locking can transparently replace the GIL
and therefore parallelise existing applications, generally without any
changes\footnote{There are rare cases where not having atomic
bytecodes actually changes the semantics.}
It does however not provide a better synchronisation
mechanism to the application like e.g. atomic blocks.

@@

While one can argue that STM requires the insertion of read and write
barriers in the whole program, this can be done automatically and
locally by a program transformation\cite{felber07}. There are attempts
to do the same for fine-grained locking\cite{bill06} but they require
a whole program analysis since locks are inherently non-composable.
The effectiveness of these approaches still has to be proven for our
use case.
[xxx or maybe: "The effectiveness of these approaches is doubtful in our
use case --- for example, it makes it close to impossible to order the
locks consistently or to know in advance which locks a transaction will
need.]
